---
title: "PracticalMachineLearning"
author: "ceeka"
date: "Saturday, January 24, 2015"
output: html_document
---
Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

Data 


The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 


```{r}
setwd("C:/Users/Karthik/Documents/R")
```


```{r}
#Importing the required libraries
library(ElemStatLearn)
library(caret)
library(rpart)
library(randomForest)
```


```{r}
#Reading the data
data <- read.csv("pml-training.csv",header=TRUE, sep=",", na.strings=c("NA",""))
dataTest <- read.csv("pml-testing.csv",header=TRUE, sep=",", na.strings=c("NA",""))
data <- data[ , colSums(is.na(data)) == 0]
```

```{r}
#Checking the dimensions of Data
dim(data)
head(data)
```

```{r}
#Removing variable that are not required
drop <- c('X','user_name','raw_timestamp_part_1','raw_timestamp_part_2','cvtd_timestamp','new_window','num_window')
data <- data[,!(names(data) %in% drop)]
```

```{r}
#Find highly correlated variable and remove
new_data <- cor(na.omit(data[sapply(data, is.numeric)]))
removeHighlyCorr = findCorrelation(new_data, cutoff = .90, verbose = TRUE)
data <- data[,-removeHighlyCorr]
dim(data)
```

```{r}
#Creating Data partition
inTrain = createDataPartition(data$classe,p=0.6,list=FALSE)
trainTree <- data[inTrain,]
dim(trainTree)
testTree <- data[-inTrain,]
```

```{r}
#CART 
#Using Tree package
library(tree)
set.seed(1232)
fit <- train(classe~.,method="rpart",data=trainTree)
```
  

```{r}
#Plotting tree
library(rattle)
fancyRpartPlot(fit$finalModel)
```

Using Tree package summarizing the confusion matrix
```{r}
tree.train = tree(classe~.,data=trainTree)
summary(tree.train)
tree.pred=predict(tree.train,testTree,type="class")
predTab = with(testTree,table(tree.pred,classe))
sum(diag(predTab))/sum(as.vector(predTab)) # error rate
```

Pruning Tree
```{r}
cv.train = cv.tree(tree.train,FUN=prune.misclass)
cv.train
plot(cv.train)
```
Pruning at size 20
```{r}
prune.train=prune.misclass(tree.train,best=10)
tree.pred=predict(prune.train,testTree,type="class")
predMatrix = with(testTree,table(tree.pred,classe))
sum(diag(predMatrix))/sum(as.vector(predMatrix)) # error rate
```

```{r}
#Using Random Forest
require(randomForest)
set.seed(123451)
rf.train <- randomForest(classe~.,data=trainTree,nTree=100,importance=TRUE)
rf.train
```

```{r}
#Plotting Rf
varImpPlot(rf.train)
importance(rf.train)
```

```{r}
#Predicting RandomForest Accuracy
rf.pred=predict(rf.train,testTree,type="class")
predRFTab = with(testTree,table(rf.pred,classe))
sum(diag(predRFTab))/sum(as.vector(predRFTab)) # error rate
```

```{r}
#Conclusion
prediction <- predict(rf.train, dataTest)
prediction
```

To Generate 20 different files of test set prediction
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(prediction)
```

```{r}
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
